Polytope notes

### 2019-12-23 notes/polytope-notes.tex
Ready for a look

\begin{definition}[Face]
	For any valid inequality of a polytope, the subset of the polytope of vectors which are tight for the inequality is called a \emph{face} of the polytope.
	That is, the set $F$ is a face of the polytope $P$ if and only if 
	\begin{align*}
	F &= \{x \in P : \inprod{a}{x} = \beta \}
	\end{align*}
	for some valid inequality $(a, \beta)$ of $P$.
\end{definition}

@@

\begin{theorem}[EPFL Thesis Theorem 3.1.2]
	Let $P_1, \ldots, P_n$ be polytopes in $\reals^d$ and let $F$ be a face of the Minkowski sum $P := P_1 \oplus \ldots \oplus P_n$.
	Then there are faces $F_1, \ldots, F_n$ of $P_1, \ldots, P_n$ respectively such that $F = F_1 \oplus \ldots \oplus F_n$.
	Moreover, this decomposition is unique.
\end{theorem}

@@

\begin{definition}[Weighted Minkowski sum]
  If $P_1, \ldots, P_n$ are polytopes in $\reals^d$, we can call $P(\vec \lambda)$ their \emph{weighted} Minkoski sum for $\vec \lambda \in \reals^n_+$
  \begin{align*}
    P(\lambda_1, \ldots, \lambda_n) = \lambda_1 P_1 \oplus \ldots \oplus \lambda_n P_n
  \end{align*}
\end{definition}

  From the thesis:  \emph{``It is easy to see that the normal fan (undefinedhere, but consequently normal cones) of $\lambda_i P_i$ does not change as long as $\lambda_i$ is positive.  Since the normal fan of a Minkowski sum can be deduced from that of its summands, we can deduce from this that the conbimatorial properties of $P(\lambda_1, \ldots, \lambda_n)$ stay the same as long as all $\lambda_i$ are positive.''}

  \section{Starting with polytopes}\label{sec:start-polytope}
  In this section, we will suppose we start with a finite set of $n$ polytopes $\T := \{T_1, \ldots, T_n\}$, and we will call $T := T_1 \oplus \ldots \oplus T_n$ their Minkowski sum.
  We know that every polytope has both a halfspace and vertex representation ($\H$-representation and $\V$ representation, respectively.)
  In order to construct the $\H$-representation, we know there must be a matrix $V$ and vector $e$ such that $T = \{x : Vx \leq e\}$, but it is unclear how to find such a set, which we discuss later.

@@

  \subsection{Finding $V$}
  \btw{Not sure if we want to go this far, or just argue that a complete set of normals exists.  This would be where Corollaries 1 and 2 come in though.}
  In order to find a $V$ that is complete with respect to $T := \oplus_y T_y$, we can look at the normal cones for each $T_i$ and find a vector in the normal cone $\N(F_y, T_y)$ in the decomposition of $F$ in $T$ for each face $F$.
  Since we know $T_1, \ldots, T_n$ are polytopes with $\H$-representations, one of the first natural steps is to find a (possibly minimal) set of valid inequalities describing each $T_i$.
  For each face $F_j$ of a given polytope $T$, we can choose a normal $v_j \in \N(F_j; T)$ and describe the face $F$ by $\{x \in T:  \inprod{v_j}{x} = \sup_{x \in T} \inprod{v_j}{x} =: e_j\}$.
  For each face $F_{j}^{y}$ that is in the decomposition of the face $F_j$, we have $\inprod {v_j}{x} \leq e^y_j$ forming a tight bound on $F_j$. 
  If we do this over each $T_y$, we can call $e^y$ the vector of such $e^y_j$ vectors for each face of $T_y$.
  We then concatenate these $e^y$ vectors to form the matrix $E$.

@@

  \begin{conjecture}
  	The set of normals $V$ and matrix $E$ constructed by taking an element of the normal cone $\N(F, T)$ for each face $F$ in $T$ is complete for both $T$ and each $T_y \in \T$.
  \end{conjecture}
  \begin{proof}
  	First, to see this constructed $V$ is complete for $T$, consider that the set of normals to the facets (in the affine hull) of $T$ is complete for $T$.
  	Moreover, each additional normal is a redundant constraint, so
%   since the inner product of elements of the normal cone is less than or equal to the support for the polytope for every element of $T$.
%  	Thus, 
  	this set of normals is complete for $T$.

  	To see why $V$ is complete with respect to each $T_y$, consider each case of subset inclusion.
  	We will let $\bar T_y$ be the polytope described by $V$ and $E$, and $T_y$ be the given polytope.

  	First,  to see $T_y \subseteq \bar T_y  := \{ x : Vx \leq E_{;y}\}$, consider some $x \in T_y$.
  	For all $i \in [k]$, we then have $\inprod{v_i}{x} \leq H_{T_y}(v_i) = e^y_i$.
  	By definition, this holds for all $v_i$ such that $v_i \in \N(F, T_y)$ for some $F$ a face of $T_y$.
  	Moreover, we want to claim this is true for all $v_i$, i.e. for all $i$, we have $v_i \in \N(F_y, T_y)$ for some face $F_y$ of $T_y$.

  	This follows from Corollary~\ref{cor:face-decomp-normal-cones}, since we know that we must have $v_i \in \N(F, T)$ for some face $F$ of $T$, and we necessarily have $v_i \in \N(F_y, T_y)$, where $F_y$ is the face of $T_y$ in the decomposition of $F$.

  	Now to see $\bar T_y \subseteq T_y$, we use a contradiction.
  	Suppose there was some $x \in \bar T_y$ such that $x \not \in T_y$.
  	Then there must be some $v_i \in V$ so that $\inprod {v_i}{x} > \sup_{y \in T_y} \inprod{v_i}{y}$.
  	This implies that for all faces $F_{j}^y$ of $T_y$, we have $v_i \not \in \N(F_{j}^y, T_y)$.
  	However, by construction of $V$, we can consider the face $F$ of $T$ such that $v_i \in \N(F,T) = \cap_y \N(F_y, T_y)$ by Corollary~\ref{cor:face-decomp-normal-cones}.
  	We necessarily have $v_i \in \N(F_y, T_y)$, yielding a contradiction.
  	Thus, we have $v_k \in \N(F_y, T_y)$, so we conclude $\bar T_y \subseteq T_y$, yielding the equality of each $\bar T_y = T_y$.
  \end{proof}

  \subsection{Once we have $V$}
  Now, for a given polytope $T(p) := \oplus p_y T_y$, we will later want to ask when a given $z \in \reals^d$ is in the polytope $T(p)$.
  We will then use this to understand the set of $p \in \simplex$ for which the given $z \in T(p)$.
  Throughout, assume we have $V$ which is complete for $\T$ and $E$ defined by the support of each normal in $V$.

@@

  As we assume $p \in \simplex$, we now describe the cell $D := \{p \in \simplex : Ep \geq \vec 0\}$ as the set of distributions such that $\vec 0 \in T(p)$.

@@

  \jessie{Notes post NeurIPS} 
  Still not sure if we can construct $B$ with the assumption that $p \in \simplex$ rather than $p \in A$, but gathering from our conversations at NeurIPS, that question seems worth investigating.
  I mean that in the sense that this $B$ matrix captures our necessary condition.
  The vertex representation seems to capture our sufficient condition, possibly rendering the simplex halfspace constraints unnecessary when we think of them as necessary conditions.

@@

  In the big picture, we want to know if we are given a cell $C$, how we can come up with a finite set of polytopes $\T = \{T_1, \ldots, T_n\}$ where $T(p) = \oplus_y p_y T_y$ such that $\vec 0 \in T(p) \iff p \in C$, or vice versa: starting from $\T$, can we derive the cell $C \subseteq \simplex$ where $\vec 0 \in T(p)$ for all $p \in C$.

@@

  \begin{lemma}\label{lem:describe-D}
    Suppose we have polytopes $\T = \{T_1, \ldots, T_n\}$ and a set of normals $V$ that is complete for $\T$.
 	Take $E = (e_{iy})$ where $e_{iy} = \max_{x \in T_y} \inprod{v_i}{x}$, and $D = \{p \in \simplex : Ep \geq \vec 0\}$.

 	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = \{p \in \simplex: Ep \geq \vec 0\}$.\btw{``stuff about $E$ and $D$.''}
  \end{lemma}
  \begin{proof}
    First, let us fix a distribution $p \in \simplex$.
    We define $T(p) := \oplus_y p_y T_y$ to be the $p$-weighted Minkowski sum over the polytopes in $\T$.
    By Theorem~\ref{thm:support-minksum}, we have the support of the (weighted) Minkowski sum is the (weighted) sum of the support of each polytope, which we can re-write as the product $Ep$.

    As each halfspace is bounded by the support by construction of $E$, then the support of the weighted polytope defined by an inequality on $v_i$ can be described as $\inprod{v_i}{z} \leq \inprod{E_i}{p}$.
    Since we this is true for all $v_i$, we then have $T(p) = \oplus_y p_y T_y = \{x \in \reals^d : Vx \leq Ep\}$.

>>>>

\begin{definition}[Face]\label{def:face}
	For any valid inequality of a polytope, the subset of the polytope of vectors which are tight for the inequality is called a \emph{face} of the polytope.
	That is, the set $F$ is a face of the polytope $P$ if and only if 
	\begin{align*}
	F &= \{x \in P : \inprod{a}{x} = \beta \}
	\end{align*}
	for some valid inequality $(a, \beta)$ of $P$.
\end{definition}

@@

\begin{theorem}[EPFL Thesis Theorem 3.1.2]\label{thm:unique-face-decomp}
	Let $P_1, \ldots, P_n$ be polytopes in $\reals^d$ and let $F$ be a face of the Minkowski sum $P := P_1 \oplus \ldots \oplus P_n$.
	Then there are faces $F_1, \ldots, F_n$ of $P_1, \ldots, P_n$ respectively such that $F = F_1 \oplus \ldots \oplus F_n$.
	Moreover, this decomposition is unique.
\end{theorem}

@@

\begin{definition}[Weighted Minkowski sum]
  If $T_1, \ldots, T_n$ are polytopes in $\reals^d$, we can call $T(\vec p)$ their \emph{weighted} Minkowski sum for $\vec p \in \reals^n_+$
  \begin{align*}
    T(\vec p) &:= \oplus_y p_y T_y = p_1 T_1 \oplus \ldots \oplus p_n T_n
  \end{align*}
\end{definition}

  From the thesis:  \emph{``It is easy to see that the normal fan (undefined here, but consequently normal cones) of $p_i T_i$ does not change as long as $p_i$ is positive.  Since the normal fan of a Minkowski sum can be deduced from that of its summands, we can deduce from this that the conbimatorial properties of $\oplus_y p_y T_y$ stay the same as long as all $p_i$ are positive.''}

  \section{Starting with polytopes}\label{sec:start-polytope}
  In this section, we will suppose we start with a finite set of $n$ polytopes $\T := \{T_1, \ldots, T_n\}$, and we will call $T := T_1 \oplus \ldots \oplus T_n \in \reals^d$ their Minkowski sum.
  We know that every polytope has both a halfspace and vertex representation ($\H$-representation and $\V$-representation, respectively.)
  In order to construct the $\H$-representation, we know there must be a matrix $V \in \reals^{k \times d}$ and vector $e \in \reals^k$ such that $T = \{x \in \reals^d : Vx \leq e\}$, but it is unclear how to find such a set, which we discuss later.

@@

  \subsection{Finding $V$}
  \btw{Not sure if we want to go this far, or just argue that a complete set of normals exists.  This would be where Corollaries 1 and 2 come in though.}
  In order to find a matrix of normals $V$ that is complete with respect to $T := \oplus_y T_y$, we can look at the normal cones for each $T_y$ and find a vector in the normal cone $\N(F_y, T_y)$ in the decomposition of $F$ in $T$ for each face $F$.
  Since we know $T_1, \ldots, T_n$ are polytopes with $\H$-representations, one of the first natural steps is to find a set of valid inequalities describing each $T_i$.
  For each face $F_j$ of the Minkowski sum polytope 
  $T$, we can choose a normal $v_j \in \N(F_j; T)$ and describe the face $F$ by $\{x \in T:  \inprod{v_j}{x} = \sup_{x \in T} \inprod{v_j}{x} =: e_j\}$ by Theorem~\ref{thm:unique-face-decomp} and Definition~\ref{def:face}.
  For each face $F_{j}^{y}$ that is in the decomposition of the face $F_j$, we have $\inprod {v_j}{x} \leq e^y_j$ forming a tight bound on $F_j$ by Theorem~\ref{thm:support-minksum}.

  For ease of notation, we concatenate the $e^y$ vectors to form the matrix $E$ according to a consistent ordering on $\Y$.

@@

  \begin{proposition}
  	The normals $V$ and matrix $E$ constructed by taking an element of the normal cone $\N(F, T)$ for each face $F$ in $T$ is complete for both $T$ and each $T_y \in \T$.
  \end{proposition}
  \begin{proof}
  	First, to see this constructed $V$ is complete for $T$, consider that the normals to the facets (in the affine span of $T$) of $T$ is complete for $T$.
  	Moreover, each additional normal is a redundant constraint, so
%   since the inner product of elements of the normal cone is less than or equal to the support for the polytope for every element of $T$.
%  	Thus, 
  	$V$ is complete for $T$.

  	To see why $V$ is complete with respect to each $T_y$, consider each case of subset inclusion.
  	We will let $\bar T_y := \{ x : Vx \leq e^y\}$ be the polytope described by $V$ and $E$, and $T_y$ be the given polytope.

  	First,  to see $T_y \subseteq \bar T_y$, take $x \in T_y$.
  	For all $i \in [k]$, we then have $\inprod{v_i}{x} \leq H_{T_y}(v_i) = e^y_i$.
  	By definition, this holds for all $v_i$ such that $v_i \in \N(F, T_y)$ for some $F$ a face of $T_y$.
  	We want to claim that, for all $v_i \in V$, there is some face $F_y$ of $T_y$ such that $v_i \in \N(F_y, T_y)$.

  	This follows from Corollary~\ref{cor:face-decomp-normal-cones}, since we know that we must have $v_i \in \N(F, T)$ for some face $F$ of $T$ by construction.
  	We necessarily have $v_i \in \N(F_y, T_y)$, where $F_y$ is the face of $T_y$ in the decomposition of $F$.

  	Now to see $\bar T_y \subseteq T_y$, we use a contradiction.
  	Suppose there was some $x \in \bar T_y$ such that $x \not \in T_y$.
  	Then there must be some $v_i \in V$ so that $\inprod {v_i}{x} > \sup_{y \in T_y} \inprod{v_i}{y}$.
  	This implies that for all faces $F_{j}^y$ of $T_y$, we have $v_i \not \in \N(F_{j}^y, T_y)$.
  	However, by construction of $V$, we can consider the face $F$ of $T$ such that $v_i \in \N(F,T) = \cap_y \N(F_y, T_y)$ by Corollary~\ref{cor:face-decomp-normal-cones}.
  	We necessarily have $v_i \in \N(F_y, T_y)$, yielding a contradiction.
  	Thus, we have $v_k \in \N(F_y, T_y)$, so we conclude $\bar T_y \subseteq T_y$, yielding the equality of each $\bar T_y = T_y$, which is the statement.
  \end{proof}

  \subsection{Once we have $V$}
  Now, for a given polytope $T(p)$, we will later want to ask when a given $z \in \reals^d$ is in the polytope $T(p)$.
  We will then use this to understand the set of $p \in \simplex$ for which the given $z \in T(p)$.
  Throughout, assume we have $V$ which is complete for $\T$ and $E$ defined by the support of each normal in $V$.

@@

  We assume $p \in \simplex$, so we now describe the cell $D^\T := \{p \in \simplex : Ep \geq \vec 0\}$ as the set of distributions such that $\vec 0 \in T(p)$.
  In cases where $\T$ is obvious from context, we forego the subscript and just write $D$.

@@

  \subsection{Notes post NeurIPS} 
  Still not sure if we can construct $B$ with the assumption that $p \in \simplex$ rather than $p \in A$, but gathering from our conversations at NeurIPS, that question seems worth investigating.
  I mean that in the sense that this $B$ matrix captures our necessary condition.
  The vertex representation seems to capture our sufficient condition, possibly rendering the simplex halfspace constraints unnecessary when we think of them as necessary conditions.
  With the results below (that nonsimplex constraints must appear in both $E$ and $B$), being able to assume the simplex for $B$ would simplify a bit... but doesn't give us a polytope construction by any means.

@@

  In the big picture, we want to know if we are given a cell $C$, how we can come up with a finite set of polytopes $\T = \{T_1, \ldots, T_n\}$ such that $\vec 0 \in T(p) \iff p \in C$, and vice versa; starting from $\T$, can we derive the cell $C \subseteq \simplex$ where $\vec 0 \in T(p)$ for all $p \in C$?

@@

  \begin{lemma}\label{lem:describe-D}
    Suppose we have polytopes $\T = \{T_1, \ldots, T_n\}$ and a set of normals $V$ that is complete for $\T$.
 	Take $E = (e_{i}^y)$ where $e_{i}^y = \max_{x \in T_y} \inprod{v_i}{x}$, and $D = \{p \in \simplex : Ep \geq \vec 0\}$.

 	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = \{p \in \simplex: Ep \geq \vec 0\}$.\btw{``stuff about $E$ and $D$.''}
  \end{lemma}
  \begin{proof}
    First, let us fix a distribution $p \in \simplex$.
    %We define $T(p) := \oplus_y p_y T_y$ to be the $p$-weighted Minkowski sum over the polytopes in $\T$.
    By Theorem~\ref{thm:support-minksum}, we have the support of the (weighted) Minkowski sum is the (weighted) sum of the support of each polytope, which we can re-write as the product $Ep$.

    As each halfspace is bounded by the support function of the weighted polytope by construction of $E$, then the support of the weighted polytope defined by an inequality on $v_i$ can be described as $\inprod{v_i}{z} \leq \inprod{E_i}{p}$.
    Taking this for all $v_i$, we then have $T(p) = \oplus_y p_y T_y = \{x \in \reals^d : Vx \leq Ep\}$.
